---
title: "ESM 244 Assignment 3"
author: "Alyssa Winchell"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1: Open Science Perspectives

##Part 2: Truckee River Flow (200 - 2016)

Attach Packages:

```{r packages}

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tseries))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(tmap))
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(ggspatial))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(raster))

```


Load Data and Convert to Time Series Data:

```{r truckeedat, message=FALSE, warning=FALSE}

truckee_flow <- read_csv("truckee_flow.csv")
truckee <- as.numeric(truckee_flow$mean_va) %>%
  na.omit()

truck_ts <- ts(truckee, frequency = 12, start = c(2000,1))

plot(truck_ts)

```

Decomposed Time Series Graph:

```{r decomp}

truck_dec <- decompose(truck_ts)
plot(truck_dec)

```

Test if Stationary:

```{r station, warning=FALSE}

adf_truck <- adf.test(truck_ts)
adf_truck

#another plot
monthplot(truck_ts)

```

The time series data appears additive because of the constant seasonal pattern and is also stationary (Dickey-Fuller test = -5.54, p= 0.01. There is a decreasing trend and seasonality as indicated by the consistent pattern and relatively large scale in the decomposition plot. Seasonal patterns occur every year while cyclical patterns occur about every 10 years.

Forcasting Truckee River Flow:
Holt Winters

```{r HW}

truck_hw <- HoltWinters(truck_ts) #performs the Holt Winters
plot(truck_hw)

truck_forecast <- forecast(truck_hw, h = 60) #forecast from HW
plot(truck_forecast, xlab = 'Year', ylab = 'Discharge (Cubic Feet Per Second', main = 'Mean Monthly Flow in the Truckee River (2000 - 2021)')

```

```{r residuals}

hist(truck_forecast$residuals)

```

The residuals appear to be normally distributed.

##Part 3: Mapping California's National Parks

Load Data: National Parks

```{r natparks, warning=FALSE}

ca_nat <- read_sf(dsn = ".", layer = "nps_boundary") %>% #read in data
  dplyr::filter(UNIT_TYPE == "National Park") #select only National Parks

st_crs(ca_nat) = 4326 #set coordinate reference system to 4326

plot(ca_nat)

```

Load Data: California Counties

```{r counties, warning=FALSE, message=FALSE}

ca_counties <- read_sf(dsn = ".", layer = "california_county_shape_file") #read in data

st_crs(ca_counties) = 4326 #set coordinate reference system to 4326

plot(ca_counties)

nat_clip <- st_intersection(ca_nat, ca_counties) #crops to just CA
plot(nat_clip)

```

Select Unique Values in Park Names for Labels:

```{r unique}

lab_name <- unique(nat_clip$UNIT_NAME) #only selects unique park names
  

```


Create a Map of National Parks:

```{r map, message=FALSE, warning=FALSE}

parks <- st_centroid(nat_clip) #finds centroids of every Park
lab_coords <- st_coordinates(parks) #finds coords for labels
nat_labs <- cbind(parks, lab_coords)

ggplot()  +
  geom_sf(data = ca_counties,
          fill = "lavender",
          color = "NA") +
  geom_sf(data= nat_clip,
          aes(fill = UNIT_NAME),
          show.legend = FALSE) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "California National Parks") +
  geom_text_repel(data = unique(nat_labs),
                  aes(X, Y, label = UNIT_NAME))


```

##Part 4: Lizards in the Northern Chihuahuan Desert – Data Wrangling and Analysis

Read in Data:

```{r lizard, message=FALSE, warning=FALSE}

desert_lizard <- read_csv("lter_lizard_pitfall.csv") %>%
  filter(site == "CALI")

```

Data Wrangling:
Do the weight of male and female lizards differ significantly?

```{r weight, message = FALSE, warning = FALSE}

mf_liz <- desert_lizard %>%
  dplyr::select(sex,weight)

female <- mf_liz %>% #data frame for only female lizards
  filter(sex == "F")

f_liz <- female$weight #pull out all the female data in the weight column. only doing this bc t test fxn has to have values instead of data frame

#create a data set called male and a vector 'm' with weights only

male <- mf_liz %>%
  filter(sex == "M")

m_liz <- male$weight

f <- as.numeric(f_liz)
m <- as.numeric(m_liz)

```

Explore for normality:

```{r}

#histogram and qqplot for f and m group
# have all 4 plots show up in one window

par(mfrow = c(2,2)) # number of rows, columns for the plots in one window
hist(f)
qqnorm(f)
hist(m)
qqnorm(m)

#shapiro-wilk test for normality test the null hyp that data are normally distributed

shapiro.test(f)
shapiro.test(m)
#in this case reject null, so it's not normal dist
#in this case, still going to use a t test for this data bc of histograms, qqplots, and bc n >> 30 in both cases

```

F test for equal variances:

```{r f_test}

# For F test:
#N0 = the ratio of sample variances equals 1 (variances are equal)
#H1: the ratio of sample variances is NOT 1 (variances not equal)

f_test <- var.test(f,m)
f_test

#with p value of 0.2938, retain the null that our sample variances are equal

```

Two-sample t tests to compare means:

```{r sigdiff}

t_test_diff <- t.test(f,m,var.equal = TRUE) # var.equal bc we est. eq var in f test
t_test_diff

```

If the mean weight of females and males is the same, the probablity that we could get means at least this different by random chance is 42.7%.

There is no significant differnce in the weights of female lizards (n = 79) and male lizards (n = 58) trapped in California (t(130) = 0.796, p = 0.427, alpha= 0.05).

Is there a significant difference in the proportion of adult female and male lizards with broken tails?

```{r tails}

tail_cond <- desert_lizard %>%
  dplyr::select(sex, tail) %>%
  dplyr::filter(sex != "J", sex != ".", tail != ".")

male_broke <- tail_cond %>%
  dplyr::filter(sex == "M", tail == "B")
male_whole <- tail_cond %>%
  dplyr::filter(sex == "M", tail == "W")

female_broke <- tail_cond %>%
  dplyr::filter(sex == "F", tail == "B")
female_whole <- tail_cond %>%
  dplyr::filter(sex == "F", tail == "W")

MT <- c(46, 10)
FT <- c(59, 18)
tail_table <- rbind(MT, FT) #makes a table of the vectors
colnames(tail_table) <- c("Whole", "Broken")
rownames(tail_table) <- c("Male", "Female")

tail_table

```

Table of Proportions:

```{r prop}

tail_prop <- prop.table(tail_table, margin = 1) #find the proportion of the data in each option by row
tail_prop

```

Hypothesis Testing:

H0: The proportiona of lizards with broken tails trapped in California is independent of sex.
HA: The proportiona of lizards with broken tails trapped in California is dependent of sex.

```{r chi}

tail_chi <-chisq.test(tail_table)
tail_chi

#p value = 0.5786

```

There is no significant association between the proportion of lizards with broken tails trapped in California and sex ($\chi$^2^(1) = 0.309, $p$ = .5786, $\alpha$ = 0.05). Notably, fewer lizards are seen with broken tail than with whole tails for both sexes. Although it is not significant, more trapped females have been found with broken tails than trapped males.
